Arrows for Parallel Computations

### 2017-12-12 paper/content/utilityFunctions.tex

Next, we also define |map|, |foldl| and |zipWith| on Arrows.
The |mapArr| combinator (Fig.~\ref{fig:mapArr}) lifts any arrow |arr a
b| to an arrow |arr [a] [b]| \cite{programming_with_arrows}. 
Similarly, we can also define |foldlArr| (Fig.~\ref{fig:foldlArr}) that lifts any arrow |arr (b, a) b| with a neutral element |b| to |arr [a] b|.

\begin{figure}[h]
\begin{code}
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
mapArr f =
	arr listcase >>>
	arr (const []) ||| (f *** mapArr f >>> arr (uncurry (:)))

listcase [] = Left ()
listcase (x:xs) = Right (x,xs)
\end{code}
\caption{The definition of |map| over Arrows and the |listcase| helper
function.}
\label{fig:mapArr}
\end{figure}

\begin{figure}[h]
\begin{code}
foldlArr :: (ArrowChoice arr, ArrowApply arr) => arr (b, a) b -> b -> arr [a] b
foldlArr f b =
	arr listcase >>>
	arr (const b) |||
		(first (arr (\a -> (b, a)) >>> f >>> arr (foldlArr f)) >>> app)
\end{code}
\caption{The definition of |foldl| over Arrows.}
\label{fig:foldlArr}
\end{figure}

@@

These combinators make use of the |ArrowChoice| type class which provides the |pipepipepipe| combinator. It takes two arrows |arr a c| and |arr b c| and combines them into a new arrow |arr (Either a b) c| which pipes all |Left a|'s to the first arrow and all |Right b|'s to the second arrow:
% \begin{figure}[htb]
\begin{code}
(pipepipepipe) :: ArrowChoice arr a c -> arr b c -> arr (Either a b) c
\end{code}
% \caption{Type signature of \pipepipepipe}
% \label{fig:codeSigPipePipePipe}
% \end{figure}

With the zipWithArr combinator we can also write a combinator |listApp|, that lifts a list of arrows |[arr a b]| to an arrow |arr [a] [b]|.
% \begin{figure}[htb]
\begin{code}
listApp :: (ArrowChoice arr, ArrowApply arr) => [arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{code}
% \caption{Definition of |listApp|}
% \label{fig:listApp}
% \end{figure}
% $ %% formatting
Note that  this additionally makes use of the |ArrowApply| type class that allows us to evaluate arrows with |app :: arr (arr a b, a) c|.

% $ %% formatting

@@

To begin with, we give the actual implementation of the non-configurable instance of the |ArrowParallel| instance:

\begin{figure}
\begin{code}
instance (NFData b, ArrowApply arr, ArrowChoice arr) => ArrowParallel arr a b () where
    parEvalN _ fs = parEvalN (hack fs) fs
                    where
                        hack :: (NFData b) => [arr a b] -> Conf b
                        hack _ = Conf rdeepseq
\end{code}
\caption{The actual default implementation of GpH's |ArrowParallel|.}
\label{fig:actualNonConfigMulticore}
\end{figure}

%
Next, we warp Eden's build-in Futures in PArrows as in
Figure~\ref{fig:RDFuture}, where |rd| is the accessor function for the |RD| wrapped inside |RemoteData|. Furthermore, in order for these |Future| types to fit with the |ArrowParallel| instances we gave earlier, we have to give the necessary |NFData| and |Trans| instances, the latter are only needed in Eden. The |Trans| instance does not have any functions declared as the default implementation suffices here. Furthermore, because |MVar| already ships with a |NFData| instance, we only have to supply a simple delegating |NFData| instance for our |RemoteData| type, where |rd| simply unwraps |RD|. The |Trans| instance does not have any functions declared as the default implementation suffices:

>>>>

Next, we give the definition of |evalN| which also helps us to define |map|, and |zipWith| on Arrows.
The |evalN| combinator in Fig.~\ref{fig:evalN} converts a list of arrows |[arr a b]| into an arrow |arr [a] [b]|.

\begin{figure}[h]
\begin{code}
evalN :: (ArrowChoice arr) => [arr a b] -> arr [a] [b]
evalN (f:fs) = arr listcase >>>
         arr (const []) ||| (f *** evalN fs >>> arr (uncurry (:)))
         where listcase []     = Left ()
               listcase (x:xs) = Right (x,xs)
evalN [] = arr (const [])
\end{code}
\caption{The definition of |evalN|}
\label{fig:evalN}
\end{figure}

The |mapArr| combinator (Fig.~\ref{fig:mapArr}) lifts any arrow |arr a
b| to an arrow |arr [a] [b]|. The original inspiration was from \cite{programming_with_arrows},
but the definition as then unified with |evalN|.

\begin{figure}[h]
\begin{code}
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
mapArr = evalN . repeat
\end{code}
\caption{The definition of |map| over Arrows.}
\label{fig:mapArr}
\end{figure}

@@

These combinators make use of the |ArrowChoice| type class which provides the |pipepipepipe| combinator. It takes two arrows |arr a c| and |arr b c| and combines them into a new arrow |arr (Either a b) c| which pipes all |Left a|'s to the first arrow and all |Right b|'s to the second arrow:
% \begin{figure}[htb]
\begin{code}
(pipepipepipe) :: ArrowChoice arr a c -> arr b c -> arr (Either a b) c
\end{code}

@@

%
We begin with warping Eden's build-in Futures in PArrows as in
Figure~\ref{fig:RDFuture}, where |rd| is the accessor function for the |RD| wrapped inside |RemoteData|. Furthermore, in order for these |Future| types to fit with the |ArrowParallel| instances we gave earlier, we have to give the necessary |NFData| and |Trans| instances, the latter are only needed in Eden. The |Trans| instance does not have any functions declared as the default implementation suffices here. Furthermore, because |MVar| already ships with a |NFData| instance, we only have to supply a simple delegating |NFData| instance for our |RemoteData| type, where |rd| simply unwraps |RD|. The |Trans| instance does not have any functions declared as the default implementation suffices:
